---
title: "R Tutorial"
date: "11/16/2023"
output: pdf_document
---

# This is a header 
## Subheader 1 
### Subheader 2   

```{r}

```


```{r}
#load packages 
library(tidyverse)
library(stargazer)
library(tinytex)

#load dataset    
whr2015 <- read_csv("~/Downloads/OneDrive - The University of Chicago/Documents/Fall 2023/Oeconomica Econometrics/WHR2015.csv") 

whr2015 <- read_csv("~/Downloads/OneDrive - The University of Chicago/Documents/Fall 2023/Oeconomica Econometrics/WHR2015.csv")
```

Shortcut to insert a code chunk is CTRL + ALT + i (Windows) or CTRL + Option + i on Mac. 

# Summary 
```{r}
summary(whr2015$`Economy (GDP per Capita)`)
```


#Filter and mutate wihtout pipes 
```{r}
filter(whr2015, Region == "Eastern Asia" | Region == "North America")

mutate(whr2015, hap_hle = `Happiness Score`*`Health (Life Expectancy)`)
```
# Pipes 
In RStudio the keyboard shortcut for the pipe operator %>% is Ctrl + Shift + M (Windows) or Cmd + Shift + M (Mac)      
```{r}

whr2015 %>% 
  group_by(Region) %>% 
  summarise(mean_gdp = median(`Economy (GDP per Capita)`))



#whr2015 %>% 
  #filter(Region == "Eastern Asia"|Region == "North America") %>% 
  #mutate(hap_hle = `Happiness Score`*`Health (Life Expectancy)`) %>% 
  #group_by(Country) %>% 
  #summarise(mean_gdp = mean(`Economy (GDP per Capita)`))



```
```{r}
#filter out missing values 
whr_cleaned <- whr2015 %>% 
  filter(!is.na(Region)) %>% 
  head(15)   
```

# Run Simple Regression Model 
```{r, results='asis'}
#creating a simple regression 

model1 <- lm(`Economy (GDP per Capita)` ~ `Health (Life Expectancy)`, data = whr2015)  

model2 <-  lm(`Economy (GDP per Capita)` ~ `Health (Life Expectancy)` + `Trust (Government Corruption)`, data = whr2015)  

#Printing regression table with stargazer 
stargazer(model1, model2, dep.var.labels = "GDP per capita", covariate.labels = c("Health", "Trust"), title = "Regression Example")

```
# Saving a dataset 
```{r}
write_csv(whr_cleaned, "~/Downloads/OneDrive - The University of Chicago/Documents/Fall 2023/Oeconomica Econometrics/whr2015new.csv")
```


# Practice 

# 1. What is the relationship between happiness score and health? Run a regression with happiness score as the independent variable and health as the dependent variable. 
# 2. Suppose you discover that freedom impacts both health and happiness score. Run a new regression with freedom as control. 
# 3. Add the dependent variable label to be "Life Expectancy as a measure of health." 
# 4. Knit the document to make sure your regression table shows. 


```{r, results='asis'}
hap_sc_health <- lm(`Health (Life Expectancy)` ~ `Happiness Score`, data = whr2015)

controlled_health <- lm(`Health (Life Expectancy)` ~ `Happiness Score` + Freedom, data = whr2015) 

stargazer(hap_sc_health, controlled_health, dep.var.labels = "Life Expectancy as a measure of health")
```

# Plots and graphs 

```{r}
whr2015 %>% 
  ggplot(mapping = aes(x = `Happiness Rank`, y = `Economy (GDP per Capita)`)) +
  geom_line() 
```






