---
title: "R Tutorial 2"
date: "12/25/2023"
output: pdf_document
---
```{r}
# Clear the environment
rm(list = ls())

#load packages 
require(tidyverse)
require(ggplot2)
require(stargazer)

# set directory
root <- "/Users/georgianaliang/Documents/GitHub/intro_to_econometrics/materials"
rawdir <- paste0(root, "/data")

#load dataset    
data <- read.csv(paste0(rawdir, "/TNeduc_cleaned.csv"))
glimpse(data)
```
## Useful Tidyverse Operations

```{r}
# filter
data %>% filter(Year == "2012")
```

```{r}
# filter by more than 1 conditions
data %>% filter(Year == "2012" & enroll_ftug < 200)
data %>% filter(Year == "2012" | enroll_ftug < 200)
data %>% filter(Year == "2012" & !(enroll_ftug < 200))
```
See the difference between "and" <&>, "or" <|>. Also the effect of "not" <!>.

```{r}
# mutate
data %>% mutate(grant_total = grant_federal + grant_state)
```

```{r}
# rename
data %>% rename(bach = degree_bach)
```

```{r}
# select
data %>% select(public:grant_state)
```

```{r}
# group_by and summarise
data %>% group_by(ID_IPEDS) %>%
  summarise(n = n(),
            avg_grant_federal = mean(grant_federal),
            avg_grant_state = mean(grant_state))
```

```{r}
# slice_
data %>% group_by(ID_IPEDS) %>%
  slice_max(grant_federal, n = 1)
```

```{r}
# rbind
data1 <- data %>% filter(Year == "2011")
data2 <- data %>% filter(Year == "2015")
data0 <- rbind(data1, data2)
glimpse(data0)
```

```{r, warning=FALSE}
# merge & join
data1 <- data %>% select(X:ID_IPEDS)
data2 <- data %>% select(ID_IPEDS:grant_state)
data0 <- merge(data1, data2, by = "ID_IPEDS")
glimpse(data0)
data00 <- inner_join(data1, data2, by = "ID_IPEDS")
glimpse(data00)
```

### Practice:
1. What is the mean total grant for public universities with 4 years or above?
2. What is the median ratio between federal and state grants for colleges with ftug less than 200?
3.* What is mean difference between grant at federal level for years 2011 and 2015

## Summary Stats
```{r, warning=FALSE, echo = FALSE, message = FALSE, echo = FALSE}
# non weighted summary stats
data %>%
  group_by(degree_bach, public) %>%
  summarise(
    min = min(grant_state),
    p10 = quantile(grant_state, probs = .1),
    p25 = quantile(grant_state, probs = .25),
    median = median(grant_state),
    p75 = quantile(grant_state, probs = .75),
    p90 = quantile(grant_state, probs = .9),
    max = max(grant_state),
    mean = mean(grant_state),
    CV = sd(grant_state)/mean(grant_state)
  )
```

```{r, warning=FALSE, echo = FALSE, message = FALSE, echo = FALSE}
require(TAM)
# weighted by ftug
data %>%
  group_by(degree_bach, public) %>%
  summarise(
    min = min(grant_state),
    p10 = weighted_quantile(grant_state, probs = .1, w = enroll_ftug),
    p25 = weighted_quantile(grant_state, probs = .25, w = enroll_ftug),
    median = weighted_quantile(grant_state, probs = .5, w = enroll_ftug),
    p75 = weighted_quantile(grant_state, probs = .75, w = enroll_ftug),
    p90 = weighted_quantile(grant_state, probs = .9, w = enroll_ftug),
    max = max(grant_state),
    mean = weighted.mean(grant_state, w = enroll_ftug/1000),
    CV = weighted_sd(grant_state, w = enroll_ftug/1000)/weighted_mean(grant_state, w = enroll_ftug/1000)
  )

```

```{r}
# correlation
cor(data$grant_federal, data$grant_state)
# weighted correlation
wtd.cors(data$grant_federal, data$grant_state, weight = data$enroll_ftug)
```

### Practice:
1. Produce a new variable that groups the colleges by number of first year first time undergraduates (ftug) enrolled in bins of 100, 200, 500, 10000, 20000, 50000. Produce a summary table for total grants at each bin level. What can you conclude about the relationship between school size (approximated by ftug number) and average grant? What about grant level variance?

## Plotting with ggplot2
```{r}
# basic plot
ggplot(data, aes(grant_state, grant_federal)) + geom_point()
```
```{r}
# varying the dots
ggplot(data, aes(grant_state, grant_federal, color = as.factor(public))) + geom_point()
ggplot(data, aes(grant_state, grant_federal, shape = as.factor(public))) + geom_point()
ggplot(data, aes(grant_state, grant_federal, size = enroll_ftug)) + geom_point()
```


```{r}
# group to different panels
ggplot(data, aes(grant_state, grant_federal)) + 
  geom_point() + 
  facet_wrap(~public+degree_bach)
```
```{r}
# add a smoothed curve
ggplot(data, aes(grant_state, grant_federal)) + 
  geom_point() + geom_smooth()
# add a linear line
ggplot(data, aes(grant_state, grant_federal)) + 
  geom_point() + geom_smooth(method = "lm")
```
```{r, , warning=FALSE}
# an alternate scatter plot using binsreg that is highly useful for studying relationships when the level of crowing of points makes it hard to think
require(binsreg)
result <- binsreg(x = data$grant_state, y = data$grant_federal)
  plot_data <- ((result$data.plot)$`Group Full Sample`)$data.dots
  p <- ggplot(aes(x = x, y = fit), data = plot_data) + geom_point() + geom_smooth(method= "lm")
```
```{r}
# add random noise for easier observation
ggplot(data, aes(as.factor(public), grant_federal)) + geom_jitter()
```

```{r}
# box plot
ggplot(data, aes(as.factor(public), grant_federal)) + geom_boxplot()
```
```{r, warning = FALSE, echo = FALSE, message = FALSE, echo = FALSE}
# histogram
ggplot(data, aes(grant_federal)) + geom_histogram()
# taking the log for clarity
ggplot(data, aes(log(grant_federal))) + geom_histogram()
```
```{r}
# bar chart
ggplot(data, aes(public)) + geom_bar()
```
```{r, warning=FALSE}
# some other features (try changing the different elements below)
  p <- ggplot(data, aes(grant_state, grant_federal, color = as.factor(public))) + geom_point() +
    scale_color_manual(name = "public/private", 
    labels = c("private", "public"), values = c("red", "green")) + 
    ylab("federal grant") + xlab("state_grant") + ggtitle("federal by state level grant") +
    xlim(c(10^5, 10^7)) + ylim(c(10^5, 6*10^6))
  p

```
```{r, echo = FALSE, message = FALSE, echo = FALSE, result = 'hide'}
# store the figure above
figdir <- paste0(root, "/figure")
ggsave(paste0(figdir, "/plot.png"), p, width = 5, height = 5)
# an alternate way of saving
output_file <- paste0(figdir, "/plot2.png")
png(output_file)
print(p)
dev.off()
```
### Practice:
1. *Create a few plots to analyze the distribution of grants by enrollment number through the years. What can you observe?

## running linear regressions
```{r}
data <- data %>% mutate(avg_grant = (grant_federal + grant_state)/enroll_ftug) %>% 
  mutate(lag_avg_grant = lag(avg_grant, n=1))
# how does enrollment change with avg grant amount
model1 <- lm(enroll_ftug ~ lag_avg_grant, data = data)
# control by public/private
model2 <- lm(enroll_ftug ~ lag_avg_grant + public, data = data)
# control by public/private +2/4 year (note the interaction)
model3 <- lm(enroll_ftug ~ lag_avg_grant + public + degree_bach + public * degree_bach, data = data)
# control by public/private +2/4 year (note the interaction) + year fixed effect
model4 <- lm(enroll_ftug ~ lag_avg_grant + public + degree_bach + public * degree_bach + as.factor(Year), data = data)

# report the information
stargazer(model1, model2, model3, model4, type = "text",
          title = "enrollment number by perivous year avg grant amount", 
          table.placement = 'H')

```

## reporting the information into latex
An efficient way to produce latex documents is to keep an organized local folder of tables and figures that we can import directly into latex. We can then compile the latex file directly without needing modify the file each time we update a specific table/graph. A great app I can recommend for this is TeXstudio.
```{r, message = FALSE}
# turn a table to latex
require(xtable)
tabdir <- paste0(root, "/tab")

# everything is reduced by 1000 times to save column space
summary_stats <- data %>%
  group_by(degree_bach, public) %>% mutate(grant_state = grant_state/1000) %>%
  summarise(
    min = min(grant_state),
    p10 = weighted_quantile(grant_state, probs = .1, w = enroll_ftug),
    p25 = weighted_quantile(grant_state, probs = .25, w = enroll_ftug),
    median = weighted_quantile(grant_state, probs = .5, w = enroll_ftug),
    p75 = weighted_quantile(grant_state, probs = .75, w = enroll_ftug),
    p90 = weighted_quantile(grant_state, probs = .9, w = enroll_ftug),
    max = max(grant_state),
    mean = weighted.mean(grant_state, w = enroll_ftug/1000),
    CV = weighted_sd(grant_state, w = enroll_ftug/1000)/weighted_mean(grant_state, w = enroll_ftug/1000)
  )
my_table <- xtable(summary_stats, digits = 2)
print.xtable(my_table, only.contents = TRUE, file = paste0(tabdir, "/tab1.tex"), include.rownames = FALSE)
```

```{r, echo=FALSE,results='hide', warning=FALSE}
# save a linear regression table
stargazer(model1, model2, model3, model4, out = paste0(tabdir, "/tab2.tex"),
          title = "enrollment number by perivous year avg grant amount", 
          table.placement = 'H', omit.stat = c("f", "ser"))
```
### Practice:
1. *Make a nice report of what you found in this tutorial using latex.
